02_DESeq2

### 2020-02-11 analysis/02_DESeq2.Rmd
adding manipulation deseq data

---
title: "06_DESeq2"
output: md_document
---

@@

source("../R/functions.R")  # load custom functions 
source("../R/themes.R")  # load custom themes and color palletes
source("../R/icons.R")  # for images

@@

# Warning: This script takes a very long time to run.

DESeq2 was not designed to run on 300 samples. But, I really like it, so I do it anyways. Some of these commands take like 15 min to run using 6 cores.

## DEseq2 on all chaacterization and manipulations

@@

>>>>

---
title: "02_DESeq2"
output: md_document
---

@@

source("../R/functions.R")  # load custom functions 
source("../R/themes.R")  # load custom themes and color palletes

@@

## Warning: This script can take a very long time to run.

DESeq2 was not designed to run on 100+ samples. But, I really like it, so I do it anyways. Some of these commands take like 15 min to run using 6 cores.

## Characterization

@@

## Manipulation

```{r manipulation}
# import "colData" which contains sample information and "countData" which contains read counts
countData <- read.csv("../results/00_counts.csv", header = T, row.names = 1)
geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1)
head(geneinfo)

# craete variable that will be critical for subset later on
colData <- read.csv("../metadata/00_samples.csv", header = T, row.names = 1)
colData$sextissue <- as.factor(paste(colData$sex, colData$tissue, sep = "_"))
colData$treatment <- factor(colData$treatment, levels = manipulationsandcontrols)
colData$tissue <- factor(colData$tissue, levels = tissuelevel)
levels(colData$treatment)
levels(colData$sex)
levels(colData$sextissue)
levels(colData$tissue)
colData <- colData %>% drop_na()

colData <- colData %>% filter(sex == "male")

createDEGdfsaveManip <- function(up, down, mytissue){

  res <- results(dds, contrast = c("treatment", up, down), independentFiltering = T, alpha = 0.1)

  DEGs <- data.frame(gene = row.names(res),
                        padj = res$padj, 
                        logpadj = -log10(res$padj),
                        lfc = res$log2FoldChange,
                        sextissue = mytissue)
  DEGs <- na.omit(DEGs)
  DEGs <- DEGs %>%
    dplyr::mutate(direction = ifelse(DEGs$lfc > 0 & DEGs$padj < 0.1, 
                                     yes = up, no = ifelse(DEGs$lfc < 0 & DEGs$padj < 0.1, 
                                                           yes = down, no = "NS"))) %>% 
    dplyr::arrange(desc(lfc))

  DEGs$direction <- factor(DEGs$direction, levels = c(down, "NS", up))

  # write DEGsframe of only significant genes
  DEGs <- DEGs %>% dplyr::filter(direction != "NS")
  print(str(DEGs))

  partialfilename = paste("_", down, "_", up, sep = "")
  myfilename = paste0("../results/DESeq2/manip/", mytissue, partialfilename, "_DEGs.csv")

  write.csv(DEGs, myfilename, row.names = F)
  # return DEGs frome with all data, included NS genes
  #print(head(DEGs))
}

for(i in levels(colData$sextissue)){

  myfilename = paste0("../results/DEseq2/manip/", i, "_vsd.csv")
  write.csv(vsd, myfilename)

  #return(dds)
  #return(vsd)
  #print(head(vsd))

  # save differential gene expression results
  inc.d3.m.inc.d3 <- createDEGdfsaveManip("m.inc.d3", "inc.d3",  i) 
  inc.d9.m.inc.d9 <- createDEGdfsaveManip("m.inc.d9", "inc.d9",  i) 
  inc.d17.m.inc.d17 <- createDEGdfsaveManip("m.inc.d17", "inc.d17",  i) 
  hatch.m.n2 <- createDEGdfsaveManip("hatch", "m.n2",  i) 
  inc.d9.m.inc.d8 <- createDEGdfsaveManip("m.inc.d8", "inc.d9",  i) 
  inc.d17.prolong <- createDEGdfsaveManip("prolong", "inc.d17",  i) 
  hatch.extend <- createDEGdfsaveManip("extend", "hatch",  i) 
}
